---
title: "Scratch"
author: "Andreas Beger, Predictive Heuristics"
date: "3/1/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library("states")
library("tidyverse")
library("haven")

source("itt.R")

data(cowstates)
cnames <- cowstates %>% group_by(cowcode) %>% summarize(country = unique(country_name)[1])

suppressMessages({
  allegations <- read_dta("data/RevisedSAFINAL_11Mar2013.dta")
  cy <- read_csv("data/CY.csv") %>%
    mutate_at(vars(starts_with("LoT")), itt_factor)
  cyvt <- read_csv("data/CYVT.csv") %>%
    mutate_at(vars(starts_with("LoT")), itt_factor)
  cyaocvt <- read_csv("data/CYAoCVT.csv")
})
```

Levels of Torture 

Ordinal scale with 6 regular values and 6 special values:

1. No Allegations
2. Infrequent
3. Several
4. Routinely
5. Widespread
6. Systematic

- Continued
- Improved
- Increased/Worsening
- Allegation, no LoT
- -777 central government collapse
- -999 foreign state occupation

```{r}
ggplot(cy, aes(x = LoT)) + 
  geom_bar() +
  coord_flip()
```

# Levels of torture by victim type

The first graph shows overall levels of torture. Unlike the CY version, it does not aggregate to highest level of torture accross victim types for a given country-year.

```{r}
cyvt %>%
  gather(key, value, starts_with("LoT")) %>%
  ggplot(aes(x = value)) +
  geom_bar() +
  coord_flip()

cyvt %>%
  gather(key, value, starts_with("LoT")) %>%
  ggplot(aes(x = value, group = key)) +
  geom_bar() +
  coord_flip() +
  facet_wrap(~ key)

cyvt %>%
  gather(key, value, starts_with("LoT")) %>%
  filter(value!="No Allegations") %>%
  ggplot(aes(x = value, group = key)) +
  geom_bar() +
  coord_flip() +
  facet_wrap(~ key)

cyvt %>%
  gather(key, value, starts_with("LoT")) %>%
  mutate(value = itt_factor(value)) %>%
  ggplot(aes(x = factor(cowccode), y = year, fill = value)) +
  facet_wrap(~ key) +
  geom_tile()
  
```

## What can be done about the non-ordinal code values?

Seven of the values that LoT can take are "regular". Five are for an ordinal scale of torture and there are 2 kinds of legitimate missing values for occuped and collapsed states. Where do the others occurr?

```{r}
cyvt_long <- cyvt %>% 
  gather(Victim, LoT, starts_with("LoT")) %>% 
  mutate(LoT = itt_factor(LoT)) %>%
  left_join(., cnames, by = c("cowccode" = "cowcode"))

cyvt_long %>% 
  filter(LoT=="-999") %>%
  group_by(cowccode) %>%
  summarize(country = unique(country), 
            years = paste0(sort(unique(year)), collapse = ";")) %>%
  knitr::kable(caption = "States under foreign occupation (-999)")

cyvt_long %>% 
  filter(LoT=="-777") %>%
  group_by(cowccode) %>%
  summarize(country = unique(country), 
            years = paste0(sort(unique(year)), collapse = ";")) %>%
  knitr::kable(caption = "Collapsed states (-777)")

cyvt_long %>% 
  filter(LoT=="Continued") %>%
  group_by(cowccode) %>%
  summarize(country = unique(country), 
            years = paste0(sort(unique(year)), collapse = ";")) %>%
  knitr::kable(caption = "LoT Continued")

cyvt_long %>% 
  filter(LoT=="Improved") %>%
  group_by(cowccode) %>%
  summarize(country = unique(country), 
            years = paste0(sort(unique(year)), collapse = ";")) %>%
  knitr::kable(caption = "LoT Improved")

cyvt_long %>% 
  filter(LoT=="Increased/Worsening") %>%
  group_by(cowccode) %>%
  summarize(country = unique(country), 
            years = paste0(sort(unique(year)), collapse = ";")) %>%
  knitr::kable(caption = "LoT Increased/Worsening")

cyvt_long %>% 
  filter(LoT=="Allegation, No LoT") %>%
  group_by(cowccode) %>%
  summarize(country = unique(country), 
            years = paste0(sort(unique(year)), collapse = ";")) %>%
  knitr::kable(caption = "Allegation, No LoT")
```


# Country-year, agency, victim

```{r}
tidy_cyaocvt <- cyaocvt %>%
  gather(starts_with("LoT"))
```

# Allegations 

```{r}
by_victim <- allegations %>%
  gather(victim, value, vtcriminal:vtunst) %>%
  filter(value==1) %>%
  mutate(value = NULL,
         victim = str_replace(victim, "vt", ""))

by_alleg <- allegations %>%
  mutate(n_vict = vtcriminal+vtdissident + vtmarginalized + vtstateagent + vtpow + vtunst) %>%
  filter(n_vict==1) %>%
  gather(victim, value, vtcriminal:vtunst) %>%
  filter(value==1) %>%
  mutate(value = NULL,
         victim = str_replace(victim, "vt", ""))

data.frame(
  victim_type = names(table(by_victim$victim)),
  by_victim_count = table(by_victim$victim) %>% as.vector(),
  by_victim_perc  = (table(by_victim$victim) / nrow(by_victim) * 100) %>% as.vector(),
  by_alleg_count = table(by_alleg$victim) %>% as.vector(),
  by_alleg_per = (table(by_alleg$victim) / nrow(by_alleg) * 100) %>% as.vector()
) %>%
  knitr::kable(digits = 1)


```

