---
title: "SI"
author: ""
date: "`r format(Sys.Date(), '%d %B %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library("dplyr")
library("tidyr")
library("readr")
```

```{r}
cy <- read_rds("../output/cy.rds")
```

For additional robustness checks, we considered the following indicators:

- Media freedom, from Whitten-Woodring and Van Belle, 2015, "The Correlates of Media Freedom", *PSRM*, via http://faculty.uml.edu/Jenifer_whittenwoodring/MediaFreedomData_000.aspx. Per the authors recommendations we collapsed the 3-valued media freedom indicator into a binary "functionally free" and "not free" version. 
- IGO membership as an indicator of political globalization, from the COW IGO state unit dataset. 
- Linear and squared year trends
- Human rights organization membership (`hro_n`) and secretariat locations (`hro_secloc`). From Murdie and Davis 2012 "Shaming and Blaming", ISQ, http://amandamurdie.org/research.html, where the original variable names are `hrfilled` and `HRsecretariatlocation`.
- Trade as % of GDP as an economic globalization indicator. From the World Bank WDI. 

The trade and HRO data contain missing values. The number of observations in a model depend on the particular combination of the 3 indicators impacted in a given specification, such that:

```{r}
cy %>%
  select(hro_n, hro_secloc, NE.TRD.GNFS.ZS) %>%
  mutate_all(is.na) %>%
  mutate(
    None = FALSE,
    `hro_n, hro_secloc` = hro_n | hro_secloc,
    `hro_n, NE.TRD.GNFS.ZS` = hro_n | NE.TRD.GNFS.ZS,
    `hro_n, hro_secloc, NE.TRD.GNFS.ZS` = hro_n | hro_secloc | NE.TRD.GNFS.ZS
    ) %>%
  summarize_all(~(n() - sum(.))) %>%
  gather(Variables, N) %>%
  mutate(Fraction = N / nrow(cy) * 100) %>%
  arrange(desc(N)) %>%
  knitr::kable(digits = 0)
```



