---
title: "Human rights orgs"
output: github_document
date: "`r format(Sys.Date(), '%d %B %Y')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library("dplyr")
library("tidyr")
library("readr")
library("ggplot2")
library("states")
library("lubridate")
```

```{r}
hro  <- read_csv("input/murdie_davis_isq_2010.csv")
cy   <- read_csv("input/cy-target-set.csv")
```

Use "hrfilled" and "HROsecretariatlocation". 

```{r}
hro <- hro %>%
  select(cowcode, year, hrfilled, HROsecretariatlocation) %>%
  rename(gwcode = cowcode) %>%
  filter(year %in% unique(cy$year))
```

Fix some country code divergences. This dataset is tricky because it tends to have all country codes, from both COW and G&W, e.g. it has 678 and 679, but one has missing values for the observations. 

```{r}
hro <- hro %>%
  # Yemen
  filter(gwcode != 678) %>%
  mutate(gwcode = ifelse(gwcode==679, 678, gwcode)) 
```


## Set overlap

Check country-year ID set overlap before merge. 

```{r}
hro_set <- hro[, c("gwcode", "year")] %>%
  mutate(in_hro = TRUE)
cy_set   <- cy %>%
  mutate(in_cy = TRUE)
full_set <- full_join(hro_set, cy_set, 
                      by = c("gwcode", "year")) %>%
  replace_na(list(in_cy = FALSE, in_hro = FALSE))

full_set %>% 
  group_by(in_hro, in_cy) %>%
  count()
```

Don't need to worry about (TRUE, FALSE), the (FALSE, TRUE) are a problem. Cases in `cy` that are missing in hro. Check which countries and years.

```{r}
full_set %>%
  filter(in_hro==FALSE & in_cy==TRUE) %>%
  group_by(gwcode) %>%
  summarize(year = paste0(year, collapse = ";"))
```

Cases are good, but there are hidden values.

```{r}
hro <- hro %>%
  mutate(in_hro = TRUE) %>%
  full_join(cy_set, by = c("gwcode", "year")) %>%
  filter(in_hro | in_cy)
```

There are missing values for both indicators, aside from the country-year set match. Check each. "hrfilled" is missing 2004 and 2005, don't consider those. 

```{r}
hro %>%
  filter(in_cy, is.na(hrfilled)) %>%
  filter(year < 2004) %>%
  group_by(gwcode) %>%
  summarize(years = paste0(year, collapse = ";")) %>%
  knitr::kable()
```

Now for "HROsecretariatlocation":

```{r}
hro %>%
  filter(in_cy, is.na(HROsecretariatlocation)) %>%
  group_by(gwcode) %>%
  summarize(years = paste0(year, collapse = ";")) %>%
  knitr::kable()
```

We can drop the excess HRO cases. 

```{r}
hro <- hro %>%
  filter(in_cy) %>%
  select(-in_hro, -in_cy) %>%
  rename(hro_n = hrfilled,
         hro_secloc = HROsecretariatlocation)
```


```{r}
write_csv(hro, "output/hro.csv")
```
