---
title: "COW IGO state unit"
output: github_document
date: "`r format(Sys.Date(), '%d %B %Y')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library("dplyr")
library("tidyr")
library("readr")
library("ggplot2")
library("states")
library("lubridate")
```

```{r}
igo  <- read_csv("input/state_year_formatv3.csv")
cy   <- read_csv("input/cy-target-set.csv")
```



```{r}
dim(igo)
dplyr::glimpse(igo[, 1:10])
range(igo$year)
range(cy$year)
```

Collapse this mother in some fashion. We can check how many IGOs a state was a member of in any given year. Another question, how many IGOs were there in each year?

First, check values:

```{r}
igo %>%
  select(-state) %>%
  gather(igo, status, -ccode, -year) %>%
  count(status)
```

-9 is missing, -1 is not independent states; 1 is full membership, 2 and 3 are associate membership and observer. 

How many IGOs per year? Consider only full membership. 

```{r}
igo_long <- igo %>%
  select(-state) %>%
  gather(igo, status, -ccode, -year) %>%
  filter(status == 1)
```

```{r}
n_igos <- igo_long %>% 
  group_by(year) %>% 
  summarize(N_IGOs = length(unique(igo)))

ggplot(n_igos, aes(x = year, y = N_IGOs)) +
  geom_line() +
  theme_minimal()
```

Collapse into number of IGOs a state was member of. 

```{r}
igo <- igo_long %>%
  group_by(ccode, year) %>%
  count(ccode, year, name = "igo_n") %>%
  left_join(n_igos, by = "year") %>%
  rename(igo_total = N_IGOs) %>%
  ungroup() %>%
  mutate(igo_p = igo_n / igo_total)
```

```{r, fig.height=6}
igo %>% 
  gather(var, value, -ccode, -year) %>% 
  mutate(year  = as.Date(sprintf("%s-01-01", year)),
         spell = id_date_sequence(year, "year"),
         ccode_id = sprintf("%s-%s", ccode, spell),
         year  = as.integer(year(year))) %>%
  ggplot(aes(x = year, y = value, group = ccode_id)) + 
  facet_wrap(~ var, scales = "free_y", ncol = 1) + 
  geom_line(alpha = .2) +
  theme_minimal()
```

```{r}
df <- igo %>% 
  filter(year > 1950) %>%
  mutate(ccode = ifelse(ccode==255 & year > 1988, 260L, ccode)) %>%
  arrange(ccode, year) %>%
  mutate(year  = as.Date(sprintf("%s-01-01", year)),
         spell = id_date_sequence(year, "year"),
         ccode_id = sprintf("%s-%s", ccode, spell),
         year  = as.integer(year(year)))

ggplot(df, aes(x = year, y = igo_n, group = ccode_id)) + 
  geom_line(alpha = .15) +
  geom_line(data = df[df$ccode %in% c(2, 260, 366), ], 
            aes(color = factor(ccode)), 
            size = 1) +
  theme_minimal() +
  scale_x_continuous(breaks = seq(1950, 2010, by = 10)) +
  labs(y = "# of IGOs member of", x = "") +
  scale_color_brewer("", type = "qual", palette = 6,
                     labels = c("2" = "USA",
                                "260" = "Germany",
                                "366" = "Estonia")) +
  theme(legend.position = "top")

ggplot(df, aes(x = year, y = igo_p, group = ccode_id)) + 
  geom_line(alpha = .15) +
  geom_line(data = df[df$ccode %in% c(2, 260, 366), ], 
            aes(color = factor(ccode)),
            size = 1) +
  theme_minimal() +
  scale_x_continuous(breaks = seq(1950, 2010, by = 10)) +
  labs(y = "fraction of IGOs member of", x = "") +
  scale_color_brewer("", type = "qual", palette = 6,
                     labels = c("2" = "USA",
                                "260" = "Germany",
                                "366" = "Estonia")) +
  theme(legend.position = "top")
```


## Set overlap

Check country-year ID set overlap before merge. 

```{r}
igo <- igo %>%
  filter(year %in% unique(cy$year))

igo_set <- igo[, c("ccode", "year")] %>%
  mutate(in_igo = TRUE)
cy_set   <- cy %>%
  mutate(in_cy = TRUE)
full_set <- full_join(igo_set, cy_set, 
                      by = c("ccode" = "gwcode", "year" = "year")) %>%
  replace_na(list(in_cy = FALSE, in_igo = FALSE))

full_set %>% 
  group_by(in_igo, in_cy) %>%
  count()
```

Don't need to worry about (TRUE, FALSE), the (FALSE, TRUE) are a problem. Cases in `cy` that are missing in IGO. Check which countries and years.

```{r}
full_set %>%
  filter(in_igo==FALSE & in_cy==TRUE) %>%
  group_by(ccode) %>%
  summarize(year = paste0(year, collapse = ";"))
```

Post-1990 Germany and Yemen. This is a standard G&W / COW divergence. Recode. 

```{r}
igo <- igo %>%
  mutate(ccode = case_when(
    ccode==255 ~ 260L,
    ccode==679 ~ 678L,
    TRUE ~ as.integer(ccode)
  )) %>%
  rename(gwcode = ccode) %>%
  mutate(in_igo = TRUE) %>%
  full_join(cy_set, by = c("gwcode", "year")) %>%
  filter(in_igo | in_cy)
```


```{r}
check <- igo %>%
  group_by(in_igo, in_cy) %>%
  count()

if (any(is.na(igo$in_igo))) {
  stop("IGO is missing some needed cases")
}
```

We can drop the excess GMFD cases. 

```{r}
igo <- igo %>%
  filter(in_cy) %>%
  select(-in_igo, -in_cy)
```


```{r}
write_csv(igo, "output/igostr.csv")
```
