---
title: "Check and clean Global Media Freedom data"
output: github_document
date: "`r format(Sys.Date(), '%d %B %Y')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library("readr")
library("dplyr")
library("tidyr")
library("ggplot2")
```

```{r}
gmfd <- read_csv("input/GMFDv3.csv")
cy   <- read_csv("input/cy-target-set.csv")
```

```{r}
dplyr::glimpse(gmfd)
range(gmfd$year)
range(cy$year)
```

We can drop the unneeded years right away before checking for country-year set overlap. 

```{r}
gmfd <- gmfd %>%
  filter(year %in% unique(cy$year))

gmfd_set <- gmfd[, c("ccode", "year")] %>%
  mutate(in_gmfd = TRUE)
cy_set   <- cy %>%
  mutate(in_cy = TRUE)
full_set <- full_join(gmfd_set, cy_set, 
                      by = c("ccode" = "gwcode", "year" = "year")) %>%
  replace_na(list(in_cy = FALSE, in_gmfd = FALSE))

full_set %>% 
  group_by(in_gmfd, in_cy) %>%
  count()
```

Don't need to worry about (TRUE, FALSE), the (FALSE, TRUE) are a problem. Cases in `cy` that are missing in GMFD. Check which countries and years.

```{r}
full_set %>%
  filter(in_gmfd==FALSE & in_cy==TRUE) %>%
  group_by(ccode) %>%
  summarize(year = paste0(year, collapse = ";"))
```

Post-1990 Germany is coded as 255 versus 260 in `cy`. 

```{r}
gmfd %>% filter(ccode==255)
```

Yemen is coded as 678 versus 679 in `cy`. 

```{r}
gmfd %>% filter(ccode==679)
```

Easy peasy. 

```{r}
gmfd <- gmfd %>%
  mutate(ccode = case_when(
    ccode==255L ~ 260L,
    ccode==679L ~ 678L,
    TRUE ~ as.integer(ccode)
  )) %>%
  rename(gwcode = ccode) %>%
  mutate(in_gmfd = TRUE) %>%
  full_join(cy_set, by = c("gwcode", "year")) %>%
  filter(in_gmfd | in_cy)
```

Check there are no missing left.

```{r}
check <- gmfd %>%
  group_by(in_gmfd, in_cy) %>%
  count()

if (any(is.na(gmfd$in_gmfd))) {
  stop("GMFD is missing some needed cases")
}
```

We can drop the excess GMFD cases. 

```{r}
gmfd <- gmfd %>%
  filter(in_cy) %>%
  select(-in_gmfd, -in_cy, -country)
```

The dataset authors recommend collapsing to a dichotomous "functinally free" version if used as IV. Do that. 

```{r}
gmfd <- gmfd %>%
  mutate(gmfd_functionallyfree = as.integer(mediascore < 3)) %>%
  select(-mediascore)
```

```{r}
write_csv(gmfd, "output/gmfd.csv")
```

